---
title: "Untitled"
author: "Zhang Yaoyuan"
date: "2021/3/14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(fpp3)
```

```{r}
data_group<- read.csv('Durable_Neworder.csv')
```

###filter thed data
```{r}
data_group1 <- data_group %>%
  mutate(month=yearmonth(DATE))%>%
  select(-DATE)
data_group1
```
```{r}
data_group2 <- data_group1 %>%
  as_tsibble(index=month)
```
```{r}
data_group2 %>%
  autoplot()
```
dcmp <- us_retail_employment %>%
  model(classical_decomposition(Employed, type = "additive")) %>% 
  components()

###classical_decomposition
```{r}
dcmp <- data_group2 %>%
  model(classical_decomposition(UMDMNO,type="multiplicative"))%>%
  components()

dcmp
```

###trend
```{r}
autoplot(data_group2, UMDMNO, color="gray")+
  autolayer(dcmp, trend, color="red")
```
###components

```{r}
dcmp %>%
  autoplot()
```
### the residuals are not that random, seasonal trend seems to be stable that is because that in this method, we assume seasonality remains unchanged
```{r}
autoplot(data_group2, UMDMNO, color="gray") +
  autolayer(dcmp, season_adjust, color = "blue") 
```
### the seasonality seem to have light effect on the overall pattern, when we use seasonal adjusted data, the trend is not smooth

### STL
```{r}
data_group2 %>%
  model(STL(UMDMNO ~ trend(window = 7) + season(window = "periodic"),
            robust = TRUE
  )) %>%
  components() ->dcmp1

dcmp1 %>%
  autoplot()
```

```{r}
data_group2 %>%
  model(STL(UMDMNO ~ trend(window = 9) + season(window = "periodic"),
            robust = TRUE
  )) %>%
  components() ->dcmp2
dcmp2 %>%
  autoplot()
```
```{r}
data_group2 %>%
  model(STL(UMDMNO ~ trend(window = 10) + season(window = "periodic"),
            robust = TRUE
  )) %>%
  components() ->dcmp3###set periodic
dcmp3 %>%
  autoplot()### white noise-residuals
```
### what happened on 2020.1?
```{r}
dcmp3 %>%
  gg_subseries(season_year)
```
### the seasonality doesn not change because we set 'periodic'

```{r}
data_group2 %>%
  model(STL(UMDMNO ~ trend(window = 13) + season(window = 8),
            robust = TRUE
  )) %>%
  components() ->dcmp4
dcmp4 %>%
  autoplot()
```
### seasonality changes, it becomes stronger and stronger


```{r}
dcmp4 %>%
  gg_subseries(season_year)
```
### ???我的妈呀这咋一点看不出来到底发生了啥


```{r}
data_group2 %>%
  model(STL(UMDMNO ~ trend(window = 13) + season(window = 10),
            robust = TRUE
  )) %>%
  components() ->dcmp5
dcmp5 %>%
  autoplot()

```
### using naive method to predict
# naive


```{r}
data_group3 <- data_group2 %>%
  filter(!is.na(UMDMNO)) 
```

```{r}
fit1 <- data_group3%>%
  model(NAIVE(UMDMNO))


fit1%>%
  forecast(h=12)%>%
  autoplot()+
  geom_point(aes(y=UMDMNO),data = data_group3,color="blue")
```
```{r}
augment(fit1)
```
```{r}
augment(fit1)%>%
  ACF(.resid)%>%
  autoplot()
```
### this is not white noise, not appropriate to do so
```{r}
SNAIVE_fit <- data_group3%>%
  model(SNAIVE(~lag("year")))

SNAIVE_fit %>%
  forecast(h=12)%>%
  autoplot()+
  geom_point(aes(y=UMDMNO),data = data_group3,colour="blue")
```
```{r}
augment(SNAIVE_fit)
```
```{r}
augment(SNAIVE_fit)%>%
  ACF(.resid)%>%
  autoplot()
```
```{r}
gg_tsresiduals(SNAIVE_fit)
```

aus_production %>%
  filter(!is.na(Bricks)) %>%
  model(RW(Bricks ~ drift())) %>%
  forecast(h = "5 years") %>%
  autoplot(aus_production, level = NULL) +
  geom_line(aes(y=Bricks),data = slice(aus_production, range(cumsum(!is.na(Bricks)))),
            linetype = "dashed", colour = "blue") +
  ggtitle("Clay brick production in Australia")
### drift
```{r}
data_group3%>%
  model(RW(UMDMNO~drift()))%>%
  forecast(h=12)%>%
  autoplot(data_group3)+
  geom_line(aes(y=UMDMNO),data = data_group3,colour="blue")
```

```{r}
fit2 <- data_group3%>%
  model(RW(UMDMNO~drift()))

fit2 %>%
  forecast(h=12)%>%
  autoplot(data_group3)+
  geom_line(aes(y=UMDMNO),data = data_group3,colour="blue")

```
```{r}
gg_tsresiduals(fit2)
```
### from above plot we can see that, the residual trem is not white noise, so the drift method may not be capable of handling such a data and doing forecasts 

```{r}
```


```{r}
final_fit <-  data_group3 %>%
  model(
    Seasonal_naive = SNAIVE(UMDMNO),
    Naive = NAIVE(UMDMNO),
    Drift = RW(UMDMNO ~ drift()),
    Mean = MEAN(UMDMNO))
```

###for all simple methods, lets check which one appears to give the best prediction
```{r}
fc <- final_fit %>%
  forecast(h=12)
```



```{r}
fc %>%
  autoplot(data_group2,level=NULL)
```

```{r}
accuracy(final_fit)
```
### the drift gives the best result

### ETS
```{r}
ETS_fit <- data_group2 %>%
  model(
    addictive= ETS(UMDMNO ~ error("A")+trend("A")+season("A")),
    multiplicative =ETS(UMDMNO ~ error("M")+trend("A")+season("M"))
  )

ETS_fc <- ETS_fit %>% forecast(h=12)

ETS_fc %>%
  autoplot(data_group2, level=NULL)
```

```{r}
components(ETS_fit) %>% autoplot()
```

###是有什么assumption吗，multiplicative，为什么season和remainder直接是直线了
```{r}
train <- data_group2 %>%
  filter(year(month) <= 2019)
test <- data_group2 %>%
  filter(year(month) >= 2020)
```


fit1 <- train %>%
  model(
    additive = ETS(Arrivals ~ error("M") + trend("A") + season("M")),
    damped = ETS(Arrivals ~ error("M") + trend("Ad") + season("M"))
  )

fc <- fit1 %>%
  forecast(h = 7)

accuracy(fc, test)

```{r}
ETS_fit <- train %>%
  model(
    addictive= ETS(UMDMNO ~ error("A")+trend("A")+season("A")),
    multiplicative =ETS(UMDMNO ~ error("M")+trend("A")+season("M"))
  )

ETS_fc <- ETS_fit %>% forecast(h=12)

ETS_fc %>%
  autoplot(data_group2, level=NULL)
```

```{r}
accuracy(ETS_fc,test)
```
addictive seems to be a better choice


