---
title: "Test-ARIMA"
output:
  html_document:
    code_folding: show
    highlight: haddock
    theme: lumen
    toc: yes
    toc_depth: 4
    toc_float: yes
  pdf_document: 
    toc: yes
    toc_depth: 4
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(fpp3)
```

```{r}
umd <- read.csv("Durable_Neworder.csv")  %>%
mutate(Date = as.Date(DATE))%>% 
mutate(Month = yearmonth(Date))%>% 
select(-DATE) %>%
as_tsibble(index = Month)
```

```{r}
umd
```

```{r}
umd %>% autoplot(UMDMNO)
```

```{r}
umd_train <- umd %>% slice(1:(n()-12))
umd_test  <- umd %>% slice((n()-11):(n()))
```

```{r}
umd_train
```

```{r}
umd_test
```


### try seasonal diff. first
```{r}
umd_train %>% autoplot(
  UMDMNO %>% difference(12)
  )    
# still not that stationary
```
### try seasonal diff. first and then regular diff.
```{r}
umd_train %>% autoplot(
  UMDMNO %>% difference(12) %>% difference(1)
)
```

### take log first
```{r}
umd_train %>% autoplot(
  log(UMDMNO) %>% difference(12) %>% difference(1)
)
# no much different
```

### KPSS test of original data
```{r}
umd_train %>%
  features(UMDMNO, unitroot_kpss)
```

the p-value is less than 0.01, indicating that the null hypothesis is rejected.   
That is, the data are not stationary. We can difference the data, and apply the KPSS test again.


### KPSS test of diff
```{r}
umd_train %>%
  mutate(diff_UMDMNO = difference(UMDMNO)) %>%
  features(diff_UMDMNO, unitroot_kpss)
```


### unitroot_ndiffs
```{r}
umd_train %>%
  features(UMDMNO, unitroot_ndiffs)
# one regular differencing is required to make it stationary.
```

### unitroot_nsdiffs
```{r}
umd_train %>%
  features(UMDMNO, unitroot_nsdiffs)
# one seasonal differencing is required to make it stationary
```







### plots of difference
```{r}
umd_train %>% mutate(diff_12 = UMDMNO %>% difference(12)) %>%
  gg_tsdisplay(diff_12, plot_type = 'partial')
```
not stationary

```{r}
umd_train %>% gg_tsdisplay(
  UMDMNO %>% difference(12) %>% difference(),
  plot_type = "partial")
```
stationary. 
so d = 1, D = 1

suggested value of p, q, P, Q: 
(1) P = 0, Q = 1
(2) P = 2, Q = 0

(3) p = 0, q = 4
(4) p = 3, q = 0




### pdq(0,1,4)+PDQ(0,1,1)
```{r}
fit_arima1 <- umd_train %>%
  model(arima = ARIMA(UMDMNO~pdq(0,1,4)+PDQ(0,1,1)))
fit_arima1 %>% report()
```

### pdq(0,1,3)+PDQ(0,1,1)
```{r}
fit_arima1 <- umd_train %>%
  model(arima = ARIMA(UMDMNO~pdq(0,1,3)+PDQ(0,1,1)))
fit_arima1 %>% report()
```

### pdq(3,1,0)+PDQ(2,1,0)
```{r}
fit_arima3 <- umd_train %>%
  model(arima = ARIMA(UMDMNO~pdq(3,1,0)+PDQ(2,1,0)))
fit_arima3 %>% report()
```


### Automatic modelling with ARIMA()
```{r}
fit_auto <- umd_train %>%
  model(arima = ARIMA(UMDMNO))
fit_auto %>% report()
```

### Automatic modelling with ARIMA() , d = 1
```{r}
fit_auto2 <- umd_train %>%
  model(arima = ARIMA(UMDMNO~pdq(d=1),stepwise = FALSE, approximation = FALSE)
                      )
fit_auto2 %>% report()
```
Best: ARIMA(2,1,3)(0,1,1)[12] 


### check the residual of the best model
```{r}
fit_auto2 %>%
  gg_tsresiduals()
```

```{r}
augment(fit_auto2) %>%
  features(.resid, ljung_box,lag = 24, dof = 6)
```
it is WN

```{r}
# accuracy train
accuracy(fit_auto2) %>%
  arrange(.model) %>%
  select(.model, .type, RMSE, MAE, MAPE, MASE)
```


```{r}
fc_ori <- fit_auto2 %>% forecast(h = 12)
fc_ori %>%
  autoplot(umd, level = NULL) +
  guides(colour=guide_legend(title="Forecast"))
```


```{r}
# accuracy test
accuracy(fc_ori, umd_test)
```






### log then diff

```{r}
umd_train %>% gg_tsdisplay(
  log(UMDMNO) %>% difference(12),
  plot_type = "partial")
```


```{r}
umd_train %>% gg_tsdisplay(
  log(UMDMNO) %>% difference(12) %>% difference(),
  plot_type = "partial")
```

stationary. 
so d = 1, D = 1
suggested value of p, q, P, Q: 
(1) P = 0, Q = 1
(1) P = 2, Q = 0

(1) p = 0, q = 4
(2) p = 3, q = 0


### pdq(0,1,4)+PDQ(0,1,1)
```{r}
fit_arima4 <- umd_train %>%
  model(arima = ARIMA(log(UMDMNO)~pdq(0,1,4)+PDQ(0,1,1)))
fit_arima4 %>% report()
```



### pdq(0,1,3)+PDQ(0,1,1)
```{r}
fit_arima6 <- umd_train %>%
  model(arima = ARIMA(log(UMDMNO)~pdq(0,1,3)+PDQ(0,1,1)))
fit_arima6 %>% report()
```

### pdq(3,1,0)+PDQ(2,1,0)
```{r}
fit_arima7 <- umd_train %>%
  model(arima = ARIMA(log(UMDMNO)~pdq(3,1,0)+PDQ(2,1,0)))
fit_arima7 %>% report()
```



### Automatic modelling with ARIMA()
```{r}
fit_auto3 <- umd_train %>%
  model(arima = ARIMA(log(UMDMNO)))
fit_auto3 %>% report()
```


```{r}
fit_auto4 <- umd_train %>%
  model(arima = ARIMA(log(UMDMNO)~pdq(d=1),stepwise = FALSE, approximation = FALSE))
fit_auto4 %>% report()
```
Model: ARIMA(0,1,5)(0,1,1)[12] 


```{r}
fit_auto4 %>% gg_tsresiduals()
```

```{r}
augment(fit_auto4) %>%
  features(.resid, ljung_box,lag = 24, dof = 6)
```
it is WN


```{r}
# accuracy
accuracy(fit_auto4) %>%
  arrange(.model) %>%
  select(.model, .type, RMSE, MAE, MAPE, MASE)
```


```{r}
fc_log <- fit_auto4 %>% forecast(h = 12)
fc_log %>%
  autoplot(umd, level = NULL) +
  guides(colour=guide_legend(title="Forecast"))
```


```{r}
accuracy(fc_log, umd_test)
```



### seasonally adjusted
```{r}
dcmp <- umd_train %>%
  model(STL(UMDMNO ~ season(window = "periodic"))) %>%
  components() %>%
  select(-.model)
```

```{r}
dcmp
```


```{r}
dcmp %>%  
  autoplot(season_year)
```



```{r}
dcmp %>% 
  autoplot(season_adjust) + xlab("Month") +
  ylab("Seasonally adjusted New Orders (Millions of Dollars)")
```

```{r}
dcmp %>% gg_tsdisplay(season_adjust, plot_type = 'partial')
```

```{r}
dcmp %>% features(season_adjust, unitroot_ndiffs)
```

```{r}
dcmp %>% features(season_adjust, unitroot_nsdiffs)
```

```{r}
dcmp %>% gg_tsdisplay(difference(season_adjust), plot_type = 'partial')
```



```{r}
fit_sa_com <- dcmp %>%
  model(arima = ARIMA(season_adjust, approximation = FALSE))
report(fit_sa_com)
```


### try pdq

```{r}
# automatic selection
fit_all <- dcmp %>%
  model(auto2 = ARIMA(season_adjust ~ pdq(d=1), 
                      stepwise = FALSE, approximation = FALSE))

```




```{r}
report(fit_all)
```

```{r}
# accuracy train
accuracy(fit_all) %>%
  arrange(.model) %>%
  select(.model, .type, RMSE, MAE, MAPE, MASE)
```

Training accuracy of seasonally adjusted ARIMA (MAPE = 3.28953) is worse than original ARIMA model (MAPE = 3.222765) and log transformed ARIMA model (MAPE = 3.214481).


```{r}
fc_sea <- fit_all %>% forecast(h = 12)
fc_sea %>%
  autoplot(dcmp, level = NULL) +
  guides(colour=guide_legend(title="Forecast"))
```


```{r}
dcmp_test <- umd_test %>%
  model(STL(UMDMNO ~ season(window = "periodic"))) %>%
  components() %>%
  select(-.model)
```


```{r}
accuracy(fc_sea, dcmp_test)
```
