---
title: "Durable_prophet"
author: "XU Zijin"
date: "20/03/2021"
output:
  pdf_document:
    toc: yes
    toc_depth: '3'
    keep_tex: yes
    latex_engine: xelatex
  html_document:
    code_folding: show
    highlight: haddock
    theme: cosmo
    toc: yes
    toc_depth: 4
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "D:/HKU/Module 4/7013 Forecasting and Predictive Analytics/Project/data")
knitr::opts_chunk$set(results =  TRUE)
```

```{r, include=FALSE}
library(tinytex)
library(dygraphs)
library(fpp3)
library(fable.prophet)
```
\pagebreak

# Prophet Model

## Read durable new order data.

```{r}
Durable_Neworder <- read.csv("Durable_Neworder.csv")
Durable_Neworder <- Durable_Neworder %>% 
  mutate(Month = yearmonth(DATE)) %>% 
  as_tsibble(index = Month) %>%
  select(Month, UMDMNO) %>%
  `colnames<-`(c("Month","Order"))

# time plot 
Durable_Neworder %>% gg_tsdisplay(Order, plot_type = "partial")
```

Split training set and test set.
```{r}
durable.train <- Durable_Neworder %>%
  filter(Month <= yearmonth("2020-01-01"))
nrow(Durable_Neworder); nrow(durable.train)
```


## Try time series forecasting.

```{r}
fit.ts.1 <- durable.train %>%
  model(prophet_1 = prophet(Order ~ 
                              growth(changepoint_prior_scale = 0.5)+
                              season(period = 12, 
                                     order = 6,
                                     type = "multiplicative",
                                     prior_scale = 10)
  ))
fc.ts.1 <- fit.ts.1 %>% forecast(h = 12)
accuracy(fit.ts.1) %>% 
  select(.model, .type, RMSE, MAPE, MASE)
accuracy(fc.ts.1, Durable_Neworder) %>% 
  select(.model, .type, RMSE, MAPE, MASE)
```


With log transformation.
```{r}
fit.ts.log1 <- durable.train %>%
  model(prophet_1 = prophet(log(Order) ~ 
                              growth(changepoint_prior_scale = 0.5)+
                              season(period = 12, 
                                     order = 6,
                                     type = "multiplicative",
                                     prior_scale = 10)
  ))
fc.ts.log1 <- fit.ts.log1 %>% forecast(h = 12)
accuracy(fit.ts.log1) %>% 
  select(.model, .type, RMSE, MAPE, MASE)
accuracy(fc.ts.log1, Durable_Neworder) %>% 
  select(.model, .type, RMSE, MAPE, MASE)
```

```{r}
augment(fit.ts.1) %>% autoplot(Order, color="gray") +
  autolayer(augment(fit.ts.1),.fitted, color="red")
fc.ts.1 %>% autoplot(Durable_Neworder%>%
                       filter(Month>=yearmonth("2015-01-01")))
```




### Tune parameters.
```{r}
# construct parameter combination dataframe
changepoint_scale <- seq(0.1,0.5, 0.1)
season_scale <- c(0.1,seq(2, 10, 2))
fourier.K <- 1:(12/2)
season.type <- c("multiplicative","additive")
para_df <- expand.grid(g_n = changepoint_scale, 
                       s_n = season_scale, 
                       K = fourier.K,
                       s.type = season.type,
                       stringsAsFactors = F)
nrow(para_df)
```

```{r}
record_df <- data.frame(Model = NA, 
                        train_MAPE = NA, 
                        test_MAPE = NA)
for (i in seq(1,nrow(para_df))){
  fit <- durable.train %>%
    model(prophet(Order ~ 
                    growth(changepoint_prior_scale = para_df[i,1])+
                    season(period = 12, 
                           order = para_df[i,3],
                           type = para_df[i,4],
                           prior_scale = para_df[i,2])
    ))
  fc <- fit %>% forecast(h = 12)
  record_df[i,1] <- i
  record_df[i,2] <- accuracy(fit) %>% select(MAPE)
  record_df[i,3] <- accuracy(fc, Durable_Neworder) %>% select(MAPE)
  if (i %% 10 ==0){
    print(i)
  }
}
```
```{r}
model_df <- as.data.frame(cbind(record_df,para_df))
head(model_df)
```

Look at the boxplots to select the parameters. 
```{r, fig.height=6, fig.width=6}
grobs <- list()
model_df %>% ggplot(aes(x = factor(s_n), y = test_MAPE, fill = s.type))+
  geom_boxplot() -> grobs[[1]]
model_df %>% ggplot(aes(x = factor(K), y = test_MAPE, fill =  s.type))+
  geom_boxplot() -> grobs[[2]]
model_df %>% ggplot(aes(x = factor(g_n), y = test_MAPE, fill = s.type))+
  geom_boxplot() -> grobs[[3]]
gridExtra::grid.arrange(grobs=grobs,ncol=1)
```

### Choose the parameter combination

* prior_scale = 4
* seasonal type = "additive"
* fourier term = 6
* changepoint_prior_scale = 0.5

```{r}
fit.ts <- durable.train %>%
  model(prophet_ts = prophet(Order ~ 
                               growth(changepoint_prior_scale = 0.5)+
                               season(period = 12, 
                                      order = 6,
                                      type = "additive",
                                      prior_scale = 4)),
        prophet_log_ts = prophet(log(Order) ~ 
                               growth(changepoint_prior_scale = 0.5)+
                               season(period = 12, 
                                      order = 6,
                                      type = "additive",
                                      prior_scale = 4)
        ))
fc.ts <- fit.ts %>% forecast(h = 12)
accuracy(fit.ts) %>% 
  select(.model, .type, RMSE, MAPE, MASE)
accuracy(fc.ts, Durable_Neworder) %>% 
  select(.model, .type, RMSE, MAPE, MASE)
```
```{r}
fit.ts %>%select(prophet_ts) %>%gg_tsresiduals()
fit.ts %>%select(prophet_log_ts) %>%gg_tsresiduals()
```

Use raw data without transformation.

## Add predictors.
Predictors are one month earlier.
```{r}
PPI <- read.csv("PPIACO.csv")
PPI <- PPI %>% mutate(Month = yearmonth(DATE)) %>% 
  as_tsibble(index = Month) %>%
  select(Month, PPIACO) %>%
  `colnames<-`(c("Month","PPI")) %>%
  filter(Month >= yearmonth("1992-01-01") & 
           Month < yearmonth("2021-01-01")) 

WeeklyHour <- read.csv("Weekly_Hours_Worked.csv")
WeeklyHour <- WeeklyHour %>% mutate(Month = yearmonth(DATE)) %>% 
  as_tsibble(index = Month) %>%
  select(Month, HOHWMN02USM065S) %>%
  `colnames<-`(c("Month","Hour")) %>%
  filter(Month >= yearmonth("1992-01-01") & 
           Month < yearmonth("2021-01-01")) 

InvSales <- read.csv("Inventory_to_Sales.csv")
InvSales <- InvSales %>% mutate(Month = yearmonth(DATE)) %>% 
  as_tsibble(index = Month) %>%
  select(Month, MNFCTRIRNSA) %>%
  `colnames<-`(c("Month","InvSalesRatio")) %>%
  filter(Month >= yearmonth("1992-01-01") & 
           Month < yearmonth("2021-01-01")) 

Durable_Neworder <- Durable_Neworder %>% 
  mutate(PPI = PPI$PPI,
         WeeklyHour = WeeklyHour$Hour,
         InvSalesRatio = InvSales$InvSalesRatio)

durable.train <- Durable_Neworder %>%
  filter(Month <= yearmonth("2020-01-01"))
```

Test set
```{r}
durable.test <- Durable_Neworder %>%
  filter(Month >= yearmonth("2020-02-01"))

durable.test <- new_data(durable.train, 12) %>%
  mutate(
    PPI = durable.test$PPI,
    WeeklyHour = durable.test$WeeklyHour,
    InvSalesRatio = durable.test$InvSalesRatio
  )
durable.test
```

```{r}
fit.x <- durable.train %>%
  model(prophet_ppi = prophet(Order ~ 
                                PPI + 
                                growth(changepoint_prior_scale = 0.5)+
                                season(period = 12, 
                                       order = 6,
                                       type = "additive",
                                       prior_scale = 4)),
        prophet_wh = prophet(Order ~ 
                               WeeklyHour +
                               growth(changepoint_prior_scale = 0.5)+
                               season(period = 12, 
                                      order = 6,
                                      type = "additive",
                                      prior_scale = 4)),
        prophet_inv = prophet(Order ~ 
                                InvSalesRatio +
                                growth(changepoint_prior_scale = 0.5)+
                                season(period = 12, 
                                       order = 6,
                                       type = "additive",
                                       prior_scale = 4)),
        prophet_ppiwh = prophet(Order ~ 
                                  PPI + WeeklyHour +
                                  growth(changepoint_prior_scale = 0.5)+
                                  season(period = 12, 
                                         order = 6,
                                         type = "additive",
                                         prior_scale = 4)),
        prophet_invwh = prophet(Order ~ 
                                  InvSalesRatio + WeeklyHour + 
                                  growth(changepoint_prior_scale = 0.5)+
                                  season(period = 12, 
                                         order = 6,
                                         type = "additive",
                                         prior_scale = 4)),
        prophet_3 = prophet(Order ~ 
                              InvSalesRatio + WeeklyHour + PPI+
                              growth(changepoint_prior_scale = 0.5)+
                              season(period = 12, 
                                     order = 6,
                                     type = "additive",
                                     prior_scale = 4))
  )
```

```{r}
accuracy(fit.x) %>% 
  select(.model, .type, RMSE, MAPE, MASE)
```

```{r}
fc.x <- fit.x %>% forecast(new_data = durable.test)
accuracy(fc.x, Durable_Neworder) %>% 
  select(.model, .type, RMSE, MAPE, MASE)
```

```{r}
augment(fit.x) %>% autoplot(Order, color="gray") +
  autolayer(augment(fit.x),.fitted, color="red")
fc.x %>% autoplot(Durable_Neworder%>%
                    filter(Month>=yearmonth("2018-01-01")),
                  level = NULL)
```
```{r}
fit.x %>% select(prophet_ppiwh) %>% gg_tsresiduals()
```


